---
title: "Physicochemical_properties"
author: "Oskar Nyberg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/PhD Work/R data work/HESTIA_Project")
knitr::opts_chunk$set(verbose = TRUE, warning = FALSE)
    library(rmarkdown)
    library(xlsx)
    library(webchem)
    library(tidyverse)

```


# PHYSICOCHEMICAL Properties 

## Wrangling operation of the QSAR Toolbox output
Close to all of the physicochemical information is queried from OECD QSAR Toolbox. The task to collect physicochemcal information for substances started from the PesticideAI.csv-file off of the HESTIA.EARTH repository. CAS numbers had SMILES configurations matched (see `CIR_query_CAS_to_SMILES.R`) and Physicochemical information was collected separately from the ecotoxicological information.

```{r Physchem_prop}

NEW_PHYSCHEM <- rbind(
  # The main dataset from OECD QSAR Toolbox
  read.csv("data/excel_references/HESTIA_Envirotox_Physchem_2023-02-08_2.csv", header = T, sep = ";"), 
  # The missing CAS.Numbers with no data.
  read.csv("data/excel_references/Missing_Physchem_CAS_DATA.csv", header = T, sep = ";")
    ) %>% 
  
  # Multi constituent substances are are filtered out. I have no good way to deal with these substances (YET!)
  #filter(Predefined.substance.type != "Multi constituent") %>% 
  # Removing fugacity models of biodegradation half-lives. not needed.
  select(-starts_with("FM")) %>% 
  # Making commas into dots for numeric conversion
  mutate_at(vars(c(8:26)), ~ as.numeric(str_replace_all(., ",", ".")),
            # For Multi constituent substances, I dont have a way to deal with the delimiters ";" across the whole dataset. NAs are being introduced across instead. I skip these substaces for now.
            #vars(c(8:27)), ~ as.numeric(sub(";[^;*$]", "", test$label)) 
            ) %>% 
  # Unit conversions to USEtox-friendly format
  mutate(log.Kow = 10^log.Kow, # Converting to L/L, removing logarithmic value
         Exp.Vapor.Pressure = 133.3*Exp.Vapor.Pressure, # Converting from mmHg to Pascal
         Selected.Vapor.Pressure = 133.3*Selected.Vapor.Pressure, # Converting from mmHg to Pascal
         Exp.Henrys.Law.Constant = 101325*Exp.Henrys.Law.Constant, # Converting from atm to Pa
         BAF..upper.trophic. = 10^BAF..upper.trophic., # Removing logarithmm
         Exp.Log.P = 10^Exp.Log.P,  # Removing logarithm
         #across(starts_with("FM"), ~ log(2)/(.x*3600))  # Converting all half-time(h) into degradation rates 1/s ###(VERY EFFICIENT OPERATION COMPARED TO PREVIOUS CONVERSIONS!!!) ###
         ) %>% 
  # Renaming columns to better fit USEtox input format data frame
  rename(Est.Kow_L.L = log.Kow,
         Exp.Kow_L.L = Exp.Log.P,
         Koc_L.kg_MCI = Koc..MCI.,
         Biodeg_BIOWIN3 = Ultimate.biodeg..Biowin.3.,
         Exp.Water.Solubility_mg.L = Exp.Water.Solubility,
         Exp.Vapor.Pressure_Pa = Exp.Vapor.Pressure,
         BAF_L.Kg = BAF..upper.trophic.,
         pKa.gain = Basic.pKa..OASIS.Regression.,
         pKa.loss = Acidic.pKa..OASIS.Regression.,
         MW.g.mol = Molecular.Weight) %>% 

  # Making a prioritization strategy, prioritizing experimental over estimated data into a new column, with an annotation column.
  mutate(
        # Biodegradation rates needs conversion from probability
        Biodeg_1s = case_when(
                               Biodeg_BIOWIN3 > 4.75 & Biodeg_BIOWIN3 < 5 ~ 4.7E-5,
                               Biodeg_BIOWIN3 > 4.25 & Biodeg_BIOWIN3 <= 4.75 ~ 6.4E-6,
                               Biodeg_BIOWIN3 > 3.75 & Biodeg_BIOWIN3 <= 4.25 ~ 3.4E-5,
                               Biodeg_BIOWIN3 > 3.25 & Biodeg_BIOWIN3 <= 3.75 ~ 9.3E-7,
                               Biodeg_BIOWIN3 > 2.75 & Biodeg_BIOWIN3 <= 3.25 ~ 5.3E-7,
                               Biodeg_BIOWIN3 > 2.25 & Biodeg_BIOWIN3 <= 2.75 ~ 2.1E-7,
                               Biodeg_BIOWIN3 > 1.75 & Biodeg_BIOWIN3 <= 2.25 ~ 1.3E-7,
                               Biodeg_BIOWIN3 <= 1.75 ~ 4.7E-5
                               ),
         # applying the division factors for extrapolating W, Sl & Sd biodegradation rates with 1:2:9 respectively, according to Usetox manual.
         KdegW = Biodeg_1s,
         KdegSl = Biodeg_1s/2,
         KdegSd = Biodeg_1s/9,
         # Degradation rates in air is different. Prioritization of OH rate constant method, taking Episuite's SUMMARY (AOP v1.92): HYDROXYL RADICALS (25 deg C) 
         KdegA = case_when(is.na(OVERALL.OH.rate.constant_cm3_molecule.sec) ~ as.numeric(NA),
                                                   #!is.na(OVERALL.OH.Half.life_d) & is.na(OVERALL.OH.rate.constant_cm3_molecule.sec) ~ log(2)/(OVERALL.OH.Half.life_d*86400),
                                                   TRUE ~ (OVERALL.OH.rate.constant_cm3_molecule.sec * 1.5E6)/2),
         Kow_L.L = case_when(is.na(Exp.Kow_L.L) ~ Est.Kow_L.L,
                             TRUE ~ Exp.Kow_L.L),
         pKaChemClass = case_when(!is.na(pKa.loss) & is.na(pKa.gain) ~ "acid",
                                 is.na(pKa.loss) & !is.na(pKa.gain) ~ "base",
                                 !is.na(pKa.loss) & !is.na(pKa.gain) ~ "amphoter",
                                 Predefined.substance.type == "Multi constituent" ~ "undefined",
                                 is.na(pKa.loss) & is.na(pKa.gain) ~ "neutral"),
         # According to the USEtox manual, The estimation procedures (regression equations) are only suitable under certain conditions. "The regressions used in USEtox for calculating the Koc for the electrolytes are suited for acids within the pKa range 0–12 and with a log Kow between -2.18 and 8.50. For bases the pKa needs to be above 2 and log Kow is between -1.66 and 7.03 (Franco & Trapp 2008)."
         Koc_L.kg_MCI = case_when(pKaChemClass == "acid" | pKaChemClass == "amphoter" & pKa.loss <12 & pKa.loss > 0 & log(Kow_L.L) >= -2.18 & log(Kow_L.L) <= 8.5  ~ Koc_L.kg_MCI,
                                  pKaChemClass == "base" | pKaChemClass == "amphoter" & pKa.gain >2 & log(Kow_L.L) >= -1.66 & log(Kow_L.L) <= 7.03 ~ Koc_L.kg_MCI,
                                  pKaChemClass == "neutral" ~ Koc_L.kg_MCI),
         Source_Kow = case_when(is.na(Est.Kow_L.L) & is.na(Exp.Kow_L.L) ~ "",
                                is.na(Exp.Kow_L.L) ~ "Estimated",
                             TRUE ~ "Experimental"),
         Vapor.Pressure_Pa = case_when(is.na(Exp.Vapor.Pressure_Pa) ~ Selected.Vapor.Pressure,
                             TRUE ~ Exp.Vapor.Pressure_Pa),
         Source_Pvap = case_when(is.na(Selected.Vapor.Pressure) & is.na(Exp.Vapor.Pressure_Pa) ~ as.character(NA),
                                 is.na(Exp.Vapor.Pressure_Pa) ~ "Estimated",
                             TRUE ~ "Experimental"),
         Sol_mg.L = case_when(is.na(Exp.Water.Solubility_mg.L) ~ Water.Solubility,
                             TRUE ~ Exp.Water.Solubility_mg.L),
         Source_Sol = case_when(is.na(Exp.Water.Solubility_mg.L) & is.na(Water.Solubility) ~ as.character(NA),
                                 is.na(Exp.Water.Solubility_mg.L) ~ "Estimated",
                             TRUE ~ "Experimental"),
         kH25C_Pa.m3.mol = case_when(is.na(Exp.Henrys.Law.Constant) ~ (Vapor.Pressure_Pa*MW.g.mol)/Sol_mg.L, 
                             TRUE ~ Exp.Henrys.Law.Constant),
         Source_KH25C = case_when(is.na(kH25C_Pa.m3.mol)  ~ as.character(NA),
                                  is.na(Exp.Henrys.Law.Constant) ~ "Calculated",
                              TRUE ~ "Experimental")
         ) %>% 
    select(CAS.Number, SMILES, Molecular.formula, Predefined.substance.type, MW.g.mol, pKaChemClass, pKa.gain, pKa.loss, Kow_L.L, Koc_L.kg_MCI, kH25C_Pa.m3.mol, Vapor.Pressure_Pa, Sol_mg.L, KdegA, KdegW, KdegSl, KdegSd, BAF_L.Kg, starts_with("Source"))


NEW_PHYSCHEM <- rbind(
  NEW_PHYSCHEM,
# Adding in the missing Physchem data for the dataset
# Dependency: 
  read.csv("data/excel_references/Missing_envirotox_data.csv", header = T, sep = ",") %>% 
    select(-X)# I have a loose file here which was interdependent on the HESTIA_HC20_DB.Rmd file. This let go of dependency.
) %>% 
    distinct(CAS.Number, .keep_all = T)


```


## Substance Properties Annotations
Creating an annotation on substances being 1. Inorganic/Organic, 2. containing Heavy metals, and 3. being Halogenated.
```{r}

# Halogenated: contains c(F, Cl, Br, I)

# Investigating smiles that contain C in the smiles (Cu, Cr, Co) yet excluding such elements with a density <5g/cm3 and Carbon (C). !

# Creatng a vector of substances that I can filter against.
# This vector is based on 3 operations: 
  # 1. Looking up all SMILES containing elements with the letter "C" and filtering out all possible carbon components
  # 2. Looking up all elements not containing carbon whatsoever
  # 3. Merging both lists as CAS.numbers to filter out.
# 332 substances are identified. Although substances without SMILES configuration will be remain in the data set!

Inorganic_vector <- rbind(NEW_PHYSCHEM %>% 
                             filter(str_detect(SMILES, paste(c("Cd", "Cs", "Co", "Cs", "Cr", "Cu", "Cl", "Ca"), collapse = "|")), 
                                     !str_detect(SMILES, paste(c("CC", "c","C+[0-9]", "C+\\(", "C+\\)", "C+\\[", "C+\\]", "C+\\#", "C+\\="), collapse="|"))),
                          NEW_PHYSCHEM %>% 
                             filter(!str_detect(SMILES, paste(c("c","C"), collapse = "|")))) %>% 
                          pull(CAS.Number)

HM_vector <- NEW_PHYSCHEM %>% # Just as in the operation above with an inorganic annotation, This creates a vector of CAS.Numbers to annotate Heavy metals presence in the SMILES configuration
            filter(str_detect(SMILES, paste(read.xlsx("data/excel_references/Elements_list.xlsx", sheetName = "Sheet1", header = T) %>% # Using a list of elements (From Wikipedia) with annotations of Elements and their density
                                                      # only looking at elements >5 g/cm3 (Ca and Cl are not counted here)
                                                      mutate(Density = as.numeric(Density)) %>% 
                                                      filter(Density >=5) %>% 
                                                      pull(Abbr), 
                                            collapse = "|"))) %>% 
              pull(CAS.Number)

Halogenated_vector <- NEW_PHYSCHEM %>% # Just as in the operation above with an inorganic annotation, This creates a vector of CAS.Numbers to annotate Halogens presence in the SMILES configuration
                        filter(str_detect(SMILES, paste(c("F", "Cl", "Br", "I"), collapse = "|"))) %>% 
                        pull(CAS.Number)

NEW_PHYSCHEM <-  NEW_PHYSCHEM %>% 
                                mutate(Substance_type = case_when(CAS.Number %in% Inorganic_vector ~ "Inorganic",
                                                                  TRUE ~ "Organic"),
                                       `Heavy Metals` = case_when(CAS.Number %in% HM_vector ~ "1", 
                                                                  TRUE ~ "0"),
                                       Halogenated = case_when(CAS.Number %in% Halogenated_vector ~ "1", 
                                                                  TRUE ~ "0" ))
rm(list = c("Inorganic_vector", "HM_vector", "Halogenated_vector"))
```

## USEPA Comptox categorized lists

Copied Chunk from Pesticide_annotations.Rmd - "Compiling lists of chemicals from the USEPA Comptox categorized lists"
Reading here to decrease dependency on the above mentioned file, which clutters the work space

```{r}
# Make a vector of file names that are to be read in upcoming function
USEPA_filenames <- list.files(path = "data/USEPA - Comptox_categories", pattern = "Chemical List", full.names = TRUE)

# Function loading all the USEPA_COMPTOX substance information annotation
# need to load 16 files with a lot of specific stuff to read in. 
USEPA_CHEMLIST <- do.call(rbind, lapply(USEPA_filenames, function(x){
  read.csv(x, header = T, sep = ",", colClasses = c("NULL", rep("character", 7), rep("numeric", 7), "character", rep("numeric", 2))) %>%
    select(CASRN, PREFERRED.NAME, IUPAC.NAME, SMILES,  MOLECULAR.FORMULA) %>%
    mutate(Source = "USEPA_COMPTOX",
           Sub_source = paste(as.character(            # Inserting part of substance information in column "Group", based on file name
             gsub("data/USEPA - Comptox_categories/Chemical List ", "",
              gsub("-2023-01-20.csv", "",
                gsub("WIKI", "Wikipedia", 
                  gsub(",.*", "", x)))))),
          Group = paste(as.character(                  # Inserting part of substance information in column "Group", based on file name
             gsub("data/USEPA - Comptox_categories/Chemical List ", "",
              gsub("-2023-01-20.csv", "",
                gsub("WIKI", "Wikipedia", 
                  sub(".*?,", "", x)))))))
            }
    )
  ) %>% 
  pivot_wider(id_cols = c(CASRN, PREFERRED.NAME, IUPAC.NAME, SMILES,  MOLECULAR.FORMULA), names_from = Sub_source, values_from = Group) %>% 
  unnest(cols = everything()) %>% 
  filter(!grepl("NOCAS", CASRN)) 

#write.xlsx(USEPA_CHEMLIST, "data/excel_references/USEPA_CHEMLIST_2023-01_20.xlsx", sheetName = "Data1", col.names = T, append = T, showNA = F)

```


# Pesticide information

```{r}
# Query the BCPC Compendium of Pesticide Common Names https://pesticidecompendium.bcpc.org
  # Query and code has been made as: 
# Pest_query <- bcpc_query(CAS_CID_list$CAS.Number, from = "cas", verbose = T)
# save.rds(Pest_query, "Pest_query.txt")
Pest_query <- readRDS("data/excel_references/Pest_query.txt")
Pest_drop.na <- Pest_query[!is.na(Pest_query)] # Removing NAs list objects

Pest_rbind <- as.data.frame(do.call(rbind, Pest_drop.na)) # Flatten the nested lists into one comprehensive data frame
Pest_rbind$Query_CAS <- rownames(Pest_rbind)

rownames(Pest_rbind) = NULL # Removing rownames

Pest_rbind[1:12] <- lapply(Pest_rbind[1:12], as.character) # Coercing all data into character. Previously each cell was defined as a list.

Pest_info <- Pest_rbind %>% # selecting the cols i want to add in the final output
                rename(CAS.Number = Query_CAS) %>% 
                select(CAS.Number, cname, iupac_name, activity, subactivity) 

 #joining the pesticide activity columns with the data frame
 NEW_PHYSCHEM <- NEW_PHYSCHEM %>% 
  left_join(., 
            Pest_info,
             by = "CAS.Number")

 rm(list = c("Pest_query", "Pest_drop.na", "Pest_rbind", "Pest_info"))
 
```

## Substance names and identifiers for both HESTIA and Envirotox
Merging all of the various data queries into one(two) outputs
```{r}
# Selected identifiers for the Envirotox database
Envirotox_subst <- readxl::read_xlsx(path = "..\\..\\Backup HESTIA\\EnviroTox DB\\envirotox_DB_20230110081208.xlsx", sheet = "substance", col_names = T, col_types = c("skip", "guess", rep("text", times = 2), rep("skip", times = 19))) %>% 
    mutate(CAS.Number = as.cas(`original CAS`)) %>% 
    select(CAS.Number, `Chemical name`, `Canonical SMILES`) 

  # Making a list of CAS.Numbers to select whether a substance is Organic or inorganic
Inorganic_vector_Enviro <- rbind(Envirotox_subst %>% 
                             filter(str_detect(`Canonical SMILES`, paste(c("Cd", "Cs", "Co", "Cs", "Cr", "Cu", "Cl", "Ca"), collapse = "|")), 
                                     !str_detect(`Canonical SMILES`, paste(c("CC", "c","C+[0-9]", "C+\\(", "C+\\)", "C+\\[", "C+\\]", "C+\\#", "C+\\="), collapse = "|"))),
                          Envirotox_subst %>% 
                             filter(!str_detect(`Canonical SMILES`, paste(c("c","C"), collapse = "|")))) %>% 
                          pull(CAS.Number)

HESTIA_Enviro_subst_type_annot <- rbind(
  # loading the HESTIA substance type annotations
  readxl::read_xlsx(path = "results/HESTIA_chem_prop_list_full.xlsx", sheet = "Slim_DB", col_names = T, col_types = c("guess", rep("text", times = 4), rep("skip", times = 3), "text")), 
  # loading and wrangling substance type annotations for the envirotox substances that are missing from the HESTIA list
  # Keep in mind that this whole operations is to have physchem data annotations for the envirotox-unique substances. it's messy and repetitive, but works!
  Envirotox_subst %>% 
    left_join(x = .,
              y = USEPA_CHEMLIST %>% 
                  rename(CAS.Number = CASRN),
              by = "CAS.Number") %>% 
    # Using the Above created "inorganic_vector_Enviro" to define inorganic/organic substance types
    mutate(Substance_type = case_when(CAS.Number %in% Inorganic_vector_Enviro ~ "Inorganic",
                                      TRUE ~ "Organic")) %>% 
    unite("Subgroup", c("DRUGBANK", "EPAPCS", "HEALTHY_BUILDING_NETWORK", "NORMAN_ITN", "OPPIN", "PPDB", "USEPA", "Wikipedia"), sep = "; ", na.rm = T, remove = T)  %>% 
    mutate(Group = case_when(
                             grepl(paste(c("Antimicrobial", "antibacterial", "antibiotic", "Antibiotics"), collapse = "|"), Subgroup, fixed = F) ~ "Antibiotic",
                             grepl("Pesticide", Subgroup) ~ "Pesticide",
                             grepl("Pharmaceutical", Subgroup) ~ "Pharmaceutical",
                             TRUE ~ case_when(Substance_type == "Organic" ~ "Other organic chemicals",
                                              is.na(`Canonical SMILES`) ~ "Unknown",
                                              TRUE ~ "Other inorganic chemicals"))
            ) %>%
      rowwise() %>% # Removing the Group category from Subgroup.
      mutate(Subgroup = case_when(grepl(paste((Group)), Subgroup) ~ gsub(paste(Group, ";", sep=""), "", Subgroup),
                                  TRUE ~ Subgroup)) %>% 
      mutate(Subgroup = case_when(grepl(paste((Group)), Subgroup) ~ gsub(paste(Group), "", Subgroup),
                                  TRUE ~ Subgroup),
             Subgroup = sub(" $", "", Subgroup),
             Subgroup = sub(";$", "", Subgroup),
             Subgroup = gsub(",", "", Subgroup),
             Subgroup = sub(" \\(s\\)", "", Subgroup),
             Subgroup = sub("^ ", "", Subgroup)
             ) %>% 
      ungroup() %>% 
      mutate_all(~na_if(., '')) %>% 
      # Warning comes in here, since many term.names are removed. Only the first is selected in the `separate` function.
      separate(col = "Chemical name", into = "term.name", sep = ";") %>% 
      mutate(term.name = case_when(!is.na(PREFERRED.NAME) ~ PREFERRED.NAME,
                              TRUE ~ term.name),
             definition = as.character(NA)) %>% 
      rename(CanonicalSMILES = `Canonical SMILES`) %>% 
      select(CAS.Number, term.name, CanonicalSMILES, Group, Subgroup, definition)
      ) %>% 
  distinct(CAS.Number, .keep_all = T)

NEW_PHYSCHEM <- NEW_PHYSCHEM %>% 
    left_join(x = ., 
              y = HESTIA_Enviro_subst_type_annot %>% 
                  select(-CanonicalSMILES),
              by = "CAS.Number")
```



```{r}
# Just want to look up molecular weights via PubChem
  source(knitr::purl("code/Pesticide_annotations.Rmd"))
mw_df <- left_join(NEW_PHYSCHEM, 
          HESTIA_Comp_info_2 %>% 
            select(CAS.Number, PubChemId), 
          by = "CAS.Number")
  
NEW_PHYSCHEM$PC_MW.g_mol <- pc_prop(mw_df$PubChemId, properties = "MolecularWeight", verbose = T)[[2]]


```

```{r}
#write.csv(NEW_PHYSCHEM, "data/PHYSCHEM_INFO.csv")
#write.xlsx(data.frame(NEW_PHYSCHEM %>% filter(is.na(SMILES)) %>% distinct(CAS.Number)), "Missing_Physchem_CAS.xlsx", sheetName = "DATA", col.names = T, row.names = F, showNA = F)

rm(list = c("Inorganic_vector_Enviro", "Envirotox_subst"))

```



